---
title: "Nearshore Network Analysis"
author: "Nick Framsted"
date: "1/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

```{r}
# importing dataset for tahoe city (closest site to the pineland site)
dat_TC <- read_csv("C:/Users/ntframst/Documents/Tahoe_nearshore_network_data/TC.txt")
#str(dat_TC)
#View(dat_TC)

# importing data for homewood site (next closest to pineland site)
dat_HW <- read_csv("C:/Users/ntframst/Documents/Tahoe_nearshore_network_data/HW.txt")
```

### plotting temperature (probeT) for TC site
```{r}
# filtering out period of NAs in ProbeT column
dat_TC %>%
  filter(TimeLocal < "2020-10-01 00:00:00") %>%
  ggplot(aes(x = TimeLocal, y = ProbeT)) +
  geom_line()

# violin plot of daily temperatures
dat_TC %>%
  mutate(day = date(TimeLocal)) %>%
  filter(!is.na(ProbeT)) %>%
  ggplot(aes(x = day, y = ProbeT, group = day)) +
  geom_violin() +
  theme_bw(base_size = 20)

# looking at daily scales of temp variability
daily_T <- dat_TC %>%
  mutate(day = date(TimeLocal)) %>%
  group_by(day) %>%
  summarize(avg_T = mean(ProbeT), min_T = min(ProbeT), max_T = max(ProbeT)) %>%
  mutate(range_T = max_T - min_T)
daily_T


# adding in season column to data, grouping rows by season to calculate seasonal average temp standard deviation
seasonal_T <- daily_T %>%
  mutate(season = ifelse(day >= "2019-12-31" & day < "2020-03-01", "W", ifelse(day >= "2020-03-01" & day < "2020-06-01", "Spr", ifelse(day >= "2020-06-01" & day < "2020-09-01", "Sum", "F")))) %>%
  filter(!is.na(range_T)) %>%
  group_by(season) %>%
  summarize(avg_daily_range = mean(range_T), max_daily_range = max(range_T))
seasonal_T


# calculating average june temperatures for planning june 2021 incubation
june_T <- dat_TC %>%
  filter(TimeLocal >= "2020-06-01 00:00:00" & TimeLocal < "2020-07-01 00:00:00") %>%
  summarize(avg_T = mean(ProbeT))

# plotting june temperatures with average june temperature overlayed
dat_TC %>%
  filter(TimeLocal >= "2020-06-01 00:00:00" & TimeLocal < "2020-07-01 00:00:00") %>%
  ggplot(aes(x = TimeLocal, y = ProbeT)) +
  geom_point() +
  geom_hline(aes(yintercept = june_T$avg_T), data = june_T, color = "red")

# plotting daily max temps for june 2020
dat_june <- dat_TC %>%
  filter(TimeLocal >= "2020-06-01 00:00:00" & TimeLocal < "2020-07-01 00:00:00") %>%
  mutate(day = day(TimeLocal)) %>%
  group_by(day) %>%
  summarize(maxT = max(ProbeT))

avg_max_june_T <- mean(dat_june$maxT)

dat_june %>%
  ggplot(aes(x = day, y = maxT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_max_june_T), color = "red")

# plot just for june 5-15th for planning for june 13th incubation
dat_june %>%
  filter(day >= 5 & day <= 15) %>%
  ggplot(aes(x = day, y = maxT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_max_june_T), color = "red")
  
```

Looks like maximum daily flucuations of temperature are on the order of 6.5C in a day and average daily fluctuations are 2-3C, thus temperature treatments should be created accordingly.

Average june 2020 temp was 12.6C, average daily max temps for june is 14.29C. The range of daily max temps from june 5-15 was a minimum of ~9.5C and a max of ~16.5C, but the average was still on par with the monthly average of 14.29C.

# Plotting Aug 2020 temps to plan for Aug. 2021 temp treatments
```{r}
# calculating average August temperatures for planning August 2021 incubation
aug_T <- dat_TC %>%
  filter(!is.na(ProbeT) & TimeLocal >= "2020-08-01 00:00:00" & TimeLocal < "2020-09-01 00:00:00") %>%
  summarize(avg_T = mean(ProbeT))
#View(aug_T)

# plotting August temperatures with average august temperature overlayed
dat_TC %>%
  filter(TimeLocal >= "2020-08-01 00:00:00" & TimeLocal < "2020-09-01 00:00:00") %>%
  ggplot(aes(x = TimeLocal, y = ProbeT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_T), data = aug_T, color = "red")

# plotting daily max temps for August 2020
dat_aug <- dat_TC %>%
  filter(!is.na(ProbeT) & TimeLocal >= "2020-08-01 00:00:00" & TimeLocal < "2020-09-01 00:00:00") %>%
  mutate(day = day(TimeLocal)) %>%
  group_by(day) %>%
  summarize(maxT = max(ProbeT))

avg_max_aug_T <- mean(dat_aug$maxT)

dat_aug %>%
  ggplot(aes(x = day, y = maxT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_max_aug_T), color = "red")

# plot just for August 5-15th for planning for August 8th incubation
dat_aug %>%
  filter(day >= 5 & day <= 15) %>%
  ggplot(aes(x = day, y = maxT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_max_aug_T), color = "red")
```

Average August 2020 temperature is ~20C, with daily fluctuations on the scale of 1-2C.

Average maximum daily temps in August 2020 were ~21C, but ranged from 20C-21.75C.

During August 5-15, 2020, the average max daily temperature was 21C and variation was very small.

# October 2020 temp analysis
```{r}
# calculating average Sept/October temperatures for planning October 2021 incubation
#View(dat_TC)
# using T variable instead of ProbeT since ProbeT has all NAs for this time period
oct_T <- dat_TC %>%
  filter(!is.na(T) & TimeLocal >= "2020-09-15 00:00:00" & TimeLocal < "2020-11-01 00:00:00") %>%
  summarize(avg_T = mean(T))
#View(oct_T)

# plotting October temperatures with average august temperature overlayed
dat_TC %>%
  filter(TimeLocal >= "2020-09-15 00:00:00" & TimeLocal < "2020-11-01 00:00:00") %>%
  ggplot(aes(x = TimeLocal, y = T)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_T), data = oct_T, color = "red")

# plotting daily max temps for October 2020
dat_oct <- dat_TC %>%
  filter(!is.na(T) & TimeLocal >= "2020-09-15 00:00:00" & TimeLocal < "2020-11-01 00:00:00") %>%
  mutate(day = day(TimeLocal)) %>%
  group_by(day) %>%
  summarize(maxT = max(T))

avg_max_oct_T <- mean(dat_oct$maxT)

dat_oct %>%
  ggplot(aes(x = day, y = maxT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_max_oct_T), color = "red")

# plot just for October 3-14th for planning for October 3-14th incubation
dat_oct %>%
  filter(day >= 3 & day <= 14) %>%
  ggplot(aes(x = day, y = maxT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_max_oct_T), color = "red")
```

Since there's a gap in data for Tahoe city water temps for the period of mid sept-october, I'm using data from Homewood site just south of Pineland instead.

# Homewood Sept-Oct Water Temps
```{r}
View(dat_HW)

dat_HW %>%
  filter(!is.na(T)) %>%
  ggplot(aes(x = TimeLocal, y = T)) +
  geom_point() +
  geom_smooth()
```

Looks like 15-17.5C range at the beginning of october at Homewood's nearshore station (based on interpolation of the data), with daily fluctuations of 1-2C in mid october.

# July-Sept Water Temp Tahoe City
```{r}
# loading in new .csv file provided by Sergio Valbuena on 9/17/21 of 2019 Tahoe City water temps for july-oct to make up for data gap for this period in 2020 data
dat_2019 <- read_csv("C:/Users/ntframst/Documents/Tahoe_nearshore_network_data/NS_TahoeCity_2019_07_01_to_2019_11_01_editedNTF.csv", skip = 1)

head(dat_2019)

# plot of mid-Sept-mid-Oct temps
dat_2019 %>%
  filter(TahoeTime > "2019-09-15 00:00:00" & TahoeTime < "2019-10-20 00:00:00") %>%
  ggplot(aes(x = TahoeTime, y = Temp_C)) +
  geom_point() +
  geom_smooth()

# calculating average Sept/October temperatures for planning October 2021 incubation
oct_T <- dat_2019 %>%
  filter(TahoeTime > "2019-10-01 00:00:00" & TahoeTime < "2019-11-01 00:00:00") %>%
  summarize(avg_T = mean(Temp_C))
#View(oct_T)

# plotting October temperatures with average oct temperature overlayed
dat_2019 %>%
  filter(TahoeTime >= "2019-10-01 00:00:00" & TahoeTime < "2019-11-01 00:00:00") %>%
  ggplot(aes(x = TahoeTime, y = Temp_C)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_T), data = oct_T, color = "red") +
  labs(title = "Average October 2019 Water Temp")

# plotting daily max temps for October 2020
dat_oct <- dat_2019 %>%
  filter(!is.na(Temp_C) & TahoeTime >= "2019-10-01 00:00:00" & TahoeTime < "2019-11-01 00:00:00") %>%
  mutate(day = day(TahoeTime)) %>%
  group_by(day) %>%
  summarize(maxT = max(Temp_C))

avg_max_oct_T <- mean(dat_oct$maxT)

dat_oct %>%
  ggplot(aes(x = day, y = maxT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_max_oct_T), color = "red") +
  labs(title = "Daily October 2019 Max Water Temp")
```

Sept-Oct temps in 2019 show that this is a period of rapid cooling in the nearshore. In early october, temps were ~13-14C with 1-2C daily temperature swings.

Average October water temp is ~12.5C, but tends to be a degree or two warmer during the first few weeks of the month. Average daily max temp is ~13.5C.

Expect temps in the field to be ~12.5-15.5C and temp treatments to be based in that range.

# November 2020 temp analysis
```{r}
# loading in new .csv file provided by Sergio Valbuena on 9/17/21 of 2019 Tahoe City water temps for july-oct to make up for data gap for this period in 2020 data
dat_2019 <- read_csv("C:/Users/ntframst/Documents/Tahoe_nearshore_network_data/NS_TahoeCity_2019_07_01_to_2019_11_01_editedNTF.csv", skip = 1)

head(dat_2019)
View(dat_2019)

# plot of nov-mid-dec temps
dat_2019 %>%
  filter(TahoeTime > "2019-11-01 00:00:00" & TahoeTime < "2019-12-15 00:00:00" & !is.na(Temp_C)) %>%
  ggplot(aes(x = TahoeTime, y = Temp_C)) +
  geom_point() +
  geom_smooth()

# calculating average november-mid-Dec temperatures for planning November 2021 incubation
nov_T <- dat_2019 %>%
  filter(TahoeTime > "2019-10-01 00:00:00" & TahoeTime < "2019-11-01 00:00:00") %>%
  summarize(avg_T = mean(Temp_C))
#View(nov_T)

# plotting Nov temperatures with average Nov temperature overlayed
dat_2019 %>%
  filter(TahoeTime >= "2019-11-01 00:00:00" & TahoeTime < "2019-12-15 00:00:00") %>%
  ggplot(aes(x = TahoeTime, y = Temp_C)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_T), data = nov_T, color = "red") +
  labs(title = "Average November 2019 Water Temp")

# plotting daily max temps for nov 2019
dat_nov <- dat_2019 %>%
  filter(!is.na(Temp_C) & TahoeTime >= "2019-11-01 00:00:00" & TahoeTime < "2019-12-01 00:00:00") %>%
  mutate(day = day(TahoeTime)) %>%
  group_by(day) %>%
  summarize(maxT = max(Temp_C))

avg_max_nov_T <- mean(dat_nov$maxT)

dat_nov %>%
  ggplot(aes(x = day, y = maxT)) +
  geom_point() +
  geom_hline(aes(yintercept = avg_max_nov_T), color = "red") +
  labs(title = "Daily Nov 2019 Max Water Temp")
```

