---
title: "211010_background_metab"
author: "Nick Framsted"
date: "1/2/2022"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# Loading in data
```{r}
dat <- read_csv("data/211010_pineland_a_e/211010_background_metab_inc.csv")


head(dat)

```

# Calculating Background NEP Rates
```{r}
# creating column of average of the 3 DO measurements for each jar
dat2 <- dat %>%
  mutate(DO_avg = (DO_1_mg_L + DO_2_mg_L + DO_3_mg_L)/3)
head(dat2)
#View(dat2)

# plotting DO data vs. time
dat2 %>%
  ggplot(aes(x = time, y = DO_avg, color = jar)) +
  geom_point() +
  geom_smooth(method = lm)
# looks like jar might be an outlier since DO increased even during the dark incubation

#### NEP ####
dat3 <- dat2 %>%
  filter(time <= hms::as.hms("10:06:00")) %>% # filtering out final DO msmnt to only get NEP
  group_by(jar) %>%
  dplyr::summarise(treatment, tank, treatment, DO_diff_mg_L = last(DO_avg) - first(DO_avg), delta_T_days = as.duration(first(time) %--% last(time))/ddays(1)) # taking difference in DO conc. between pre- and post-light exposure and difference in time (in days) to calculate NEP rates

#View(dat3)

# now converting DO concentrations in mg/L to DO mass in mg by multiplying by jar volumes
dat3 <- dat3 %>%
  mutate(DO_avg_mass_mg = DO_diff_mg_L * 0.473)

# calculating background NEP rates (in mg O2 per day) for the ambient and enriched water(NEP rates for jars are multiplied by the ratio of water volume of the periphyton chambers:jars to account for the difference in water volume)
# one mason jar is 0.473L, each periphyton chamber is ~2.75L with a rock inside
dat3 <- dat3 %>%
  mutate(NEP = DO_avg_mass_mg/delta_T_days * (2.75/.473))
View(dat3)
```

Background NEP rates range from -8 to 0 mgO2 O2/day which is much smaller than the rates produced by the periphyton (range: 41-130 mgO2/day; avg: 91).

# Calculating background ER rates
```{r}
# calculating ER rates in jars
dat_ER <- dat2 %>%
  filter(time >= hms::as.hms("09:57:00")) %>% # filtering out initial DO msmnt to only get NEP
  group_by(jar) %>%
  summarise(treatment, tank, treatment, DO_diff_mg_L = first(DO_avg) - last(DO_avg), delta_T_days = as.duration(first(time) %--% last(time))/ddays(1)) # taking difference in DO conc. between pre- and post-dark exposure and difference in time (in days) to calculate ER rates

#View(dat_ER)

# now converting DO concentrations in mg/L to DO mass in mg by multiplying by jar volumes
dat_ER <- dat_ER %>%
  mutate(DO_avg_mass_mg = DO_diff_mg_L * 0.473)

# calculating background ER rates (in mg O2 per day) for the ambient and enriched water(ER rates for jars are multiplied by the ratio of water volume of the periphyton chambers:jars to account for the difference in water volume)
# one mason jar is 0.473L, each periphyton chamber is ~2.75L with a rock inside
dat_ER <- dat_ER %>%
  mutate(ER = DO_avg_mass_mg/delta_T_days * (2.75/.473))
View(dat_ER)
```

Background ER rates range from -3 to 2mg O2/day per chamber, which is much smaller than the rates produced by the periphyton (range: 9-44mg O2/day; average: 26mg O2/day). These values are perhaps large enough that they should be accounted for in periphyton ER rate determination.