---
title: "Main Periphyton Analysis"
author: "Nick Framsted"
date: "7/29/2021"
output: pdf_document
---
Analysis of all Tahoe Periphyton Metabolism Data

Input data files: Individual dataframes generated from separate analyses of each incubation.

Output data files: Main dataframe of combined dataframes from each incubation.

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(plotly)
library(reshape2)
library(DescTools)
library(arm)
library(nlme)
library(lattice)

# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# Loading and formatting dataframe
```{r, include = FALSE}
# importing, cleaning up, and combining metabolism data from all incubations
### 210301-210315 TC Ambient & Enriched data
dat1 <- read_csv("data/main_analysis/210301_210315_TC_A_E_rates.csv")


### 210407-210421 Pineland A & E data
dat2 <- read_csv("data/main_analysis/210407_210421_pnld_A_E_rates.csv")


### 210613 Pineland A & E data

dat3 <- read_csv("data/main_analysis/210613_pnld_A_E_rates.csv")

### 210808 Pineland A & E data

dat4 <- read_csv("data/main_analysis/210808_pnld_A_E_rates.csv")


### 211010 Pineland A & E data

dat5 <- read_csv("data/main_analysis/211010_pnld_A_E_rates.csv")


### 211128 Pineland A & E data
dat6 <- read_csv("data/main_analysis/211128_pnld_A_E_rates.csv")

### 220227 Pineland A & E data
dat7 <- read_csv("data/main_analysis/220227_pnld_A_E_rates.csv")



########## Data Formatting #############

dat1 <- dat1 %>%
  mutate(season = "spring", month = "march") # adding season and month columns to dataframe



dat2 <- dat2 %>%
  mutate(season = "spring", month = "april")


dat3 <- dat3 %>%
  mutate(season = "summer", month = "june")


dat4 <- dat4 %>%
  mutate(season = "summer", month = "august")

dat5 <- dat5 %>%
  mutate(season = "fall", month = "october")

dat6 <- dat6 %>%
  mutate(season = "winter", month = "november")

dat7 <- dat7 %>%
  mutate(season = "spring", month = "february")

######### Combining dataframes #########
dat <- bind_rows(dat1, dat2, dat3, dat4, dat5, dat6, dat7)
head(dat)

# adding in raw nutrient concentrations of ambient and enriched water for each experiment
dat <- dat %>%
  mutate(NO3_N_ug_L = case_when(
    nutrient_trt == "ambient" & month == "march" ~ 1.41,
    nutrient_trt == "enriched" & month == "march" ~ 534.89,
    nutrient_trt == "ambient" & month == "april" ~ 7.03,
    nutrient_trt == "enriched" & month == "april" ~ 45.19,
    nutrient_trt == "ambient" & month == "june" ~ 11.01,
    nutrient_trt == "enriched" & month == "june" ~ 374.8,
    nutrient_trt == "ambient" & month == "august" ~ 7.5,
    nutrient_trt == "enriched" & month == "august" ~ 387.33,
    nutrient_trt == "ambient" & month == "october" ~ 4.53,
    nutrient_trt == "enriched" & month == "october" ~ 376.76
  ),
  NH4_N_ug_L = case_when(
    nutrient_trt == "ambient" & month == "march" ~ 0.39,
    nutrient_trt == "enriched" & month == "march" ~ 571.52,
    nutrient_trt == "ambient" & month == "april" ~ 1.22,
    nutrient_trt == "enriched" & month == "april" ~ 45.31,
    nutrient_trt == "ambient" & month == "june" ~ 0, # if concentrations were negative, I entered them in as 0
    nutrient_trt == "enriched" & month == "june" ~ 395.58,
    nutrient_trt == "ambient" & month == "august" ~ 4.57,
    nutrient_trt == "enriched" & month == "august" ~ 416.04,
    nutrient_trt == "ambient" & month == "october" ~ 5.03,
    nutrient_trt == "enriched" & month == "october" ~ 438.69
  ),
  SRP_ug_L = case_when(
    nutrient_trt == "ambient" & month == "march" ~ 2.77,
    nutrient_trt == "enriched" & month == "march" ~ 823.85,
    nutrient_trt == "ambient" & month == "april" ~ 2.53,
    nutrient_trt == "enriched" & month == "april" ~ 67.91,
    nutrient_trt == "ambient" & month == "june" ~ 3.17,
    nutrient_trt == "enriched" & month == "june" ~ 315.47,
    nutrient_trt == "ambient" & month == "august" ~ 5.43,
    nutrient_trt == "enriched" & month == "august" ~ 320.3,
    nutrient_trt == "ambient" & month == "october" ~ 3.41,
    nutrient_trt == "enriched" & month == "october" ~ 325.74
  ))


####### Adding in Unique Chamber ID for each Experiment #########
dat <- dat %>%
  mutate(exp_ID = case_when(
    nutrients == 0 & month == "march" ~ 1,
    nutrients == 1 & month == "march" ~ 2, # since these were separate experiments on different periphyton samples, they get unique experiment IDs
    nutrients == 0 & month == "april" ~ 3,
    nutrients == 1 & month == "april" ~ 4,
    month == "june" ~ 5, # june, aug, and oct experiments used same rocks and chamber assignments for ambient and enriched runs, so it gets the same experiment ID
    month == "august" ~ 6,
    month == "october" ~ 7,
    month == "november" ~ 8,
    month == "february" ~ 9
  ))

# Creating rock_ID column
# using chamber and experiment # columns to give a unique code for each individual rock (this will indicate some sort of chamber or rock effect from repeated measures during the ambient and enriched nutrient runs)

dat <- dat %>%
  mutate(rock_ID = case_when(
    exp_ID == 5 & nutrients == 1 ~ row_number() - 16, # accouting for repeated measures on same rocks for the 5,6, and 7th experiments
    exp_ID == 6 & nutrients == 0 ~ row_number() - 16,
    exp_ID == 6 & nutrients == 1 ~ row_number() - 32,
    exp_ID == 7 & nutrients == 0 ~ row_number() - 32,
    exp_ID == 7 & nutrients == 1 ~ row_number() - 48,
    exp_ID == 8 & nutrients == 0 ~ row_number() - 48,
    exp_ID == 8 & nutrients == 1 ~ row_number() - 64,
    exp_ID == 9 & nutrients == 0 ~ row_number() - 64,
    exp_ID == 9 & nutrients == 1 ~ row_number() - 80,
     TRUE ~ as.double(row_number())
    ))
#View(dat)

# making rock_ID column a factor instead of numeric for analysis
dat$rock_ID <- as.factor(dat$rock_ID)



```


# Adding in Chla data
```{r}
####### Adding chla data by rock_ID ############

chla_main <- read_csv("data/main_analysis/chla_main_data.csv")

# formatting date column to be a date instead of a character class
chla_main$date <- mdy(chla_main$date)
head(chla_main)

# Creating rock_ID column
# using chamber and experiment # columns to give a unique code for each individual rock (this will indicate some sort of chamber or rock effect from repeated measures during the ambient and enriched nutrient runs)

chla_main <- chla_main %>%
  mutate(rock_ID = as.double(row_number())
    )
#View(chla_main)

# making rock_ID column a factor instead of numeric for analysis
chla_main$rock_ID <- as.factor(chla_main$rock_ID)

# trimming off chamber column so there aren't duplicate columns in the main dataframe

chla_main <- chla_main %>%
  dplyr::select(total_chla_ug, date, rock_ID)

# combining chla and main data by rock ID

dat <- left_join(dat, chla_main, by = "rock_ID")
#View(dat)

```


# Adding in corrected AFDW data
```{r}
######### Calculating Corrected AFDW data ###############

# since PES drying oven was found to be not operating at correct temperatures after the 11/28/21 experiment, all AFDW samples up to this point had to be redone to ensure they were ashed at proper temperatures

# As a result, we remeasured AFDW of all samples using left-over frozen sample; however, wet wts were found to no longer be accurate due to slight decrease in the water content of samples that had been frozen.

# Since we could no longer scale up these re-measured AFDW samples by wet weight, we chose to scale them up by dry weight. We were able to do this because dry weights from the first set of AFDW samples were still accurate, allowing us to calulate total Dry wt of samples by scaling up using the recorded wet wts of the fresh samples.

# We then use these total dry wts calculated by the first set of AFDW samples to scale up the re-measured AFDW samples using their recorded dry wts (SDW_60C_g).

# These corrected AFDW samples will be stored in a column titled "total_AFDW_g" and will replace the previous column of incorrect AFDW data of the same name.

# loading in re-measured AFDW data
afdw_main <- read_csv("data/main_analysis/AFDW_main_dat.csv", skip = 1)


# Creating rock_ID column
# using chamber and experiment # columns to give a unique code for each individual rock (this will indicate some sort of chamber or rock effect from repeated measures during the ambient and enriched nutrient runs)

afdw_main <- afdw_main %>%
  mutate(rock_ID = as.double(row_number() + 16)) # adding 16 to rock_ID to account for lack of re-measured afdw data for the 3/1/21 experiment in this dataset
         
#View(afdw_main)

# making rock_ID column a factor instead of numeric for analysis
afdw_main$rock_ID <- as.factor(afdw_main$rock_ID)

# selecting only columns I need

afdw_main <- afdw_main %>%
  dplyr::select(SDW_60C_g, SCW_500C_g, rock_ID)

# combining afdw and main data by rock ID column

dat <- left_join(dat, afdw_main, by = "rock_ID")

# calculating corrected AFDW data and recalculating AFDW-normalized metabolic rates
dat <- dat %>%
  mutate(total_AFDW_g_2 = case_when(
    exp_ID == 1 ~ total_AFDW_g, # using old AFDW data since there was no leftover sample
    TRUE ~ (total_DW_g/SDW_60C_g) * (SDW_60C_g - SCW_500C_g)
    )) %>% # recalculating AFDW using new data and scaling up based on previous estimates of total Dry weight
  mutate(GPP_mgO2_d_gAFDW_2 = GPP_mg_d/total_AFDW_g_2, NEP_mgO2_d_gAFDW_2 = NEP_mg_d/total_AFDW_g_2, ER_mgO2_d_gAFDW_2 = ER_mg_d/total_AFDW_g_2) # using corrected AFDW data to calculate AFDW-normalized metabolic rates

#View(dat)
   


```


# Exporting and Cleaning data
```{r}
# exporting combined dataframe as .csv
write_csv(dat, "data/main_analysis/main_rate_data.csv")

#View(dat)

# removing data from march experiments since it was at a different site and had excessive variation in AFDW-normalized GPP rates due to a discrepancy in filtering methods
dat_pnld <- dat %>%
  filter(!month == "march")
#View(dat_pnld)
glimpse(dat_pnld)

# Chambers 1 and 3 appear to be outliers for the june ambient incubations, and will be removed in subsequent analyses.

# removing outliers
# chambers 15 and 2 in Nov 2021 experiments had unusually high values for both the ambient and enriched experiments and were thus removed from the analysis

dat_clean <- dat_pnld %>%
  filter(month != "june" | nutrients != 0 | chamber != 1) %>%
  filter(month != "june" | nutrients != 0 | chamber != 3) %>%
  filter(month != "november" | chamber != 2) %>%
  filter(month != "november" | chamber != 15)
#View(dat_clean)
```


# Plotting main data
## Comparing AFDW: original vs. corrected
```{r}
AFDW_plot <- dat_clean %>%
  ggplot(aes(x = total_AFDW_g, y = total_AFDW_g_2)) +
  geom_point() +
  geom_abline(aes(intercept = 0, slope = 1)) +
  geom_rug(sides = "bl") +
  labs(x = "Old AFDW", y = "Corrected AFDW", title = "Comparison of AFDW Data") +
  lims( x = c(0, 1.5), y = c(0, 1.5)) +
  theme_classic()
AFDW_plot

ggsave("plots/main_analysis/AFDW_comparison.png", plot = AFDW_plot, width = 3, height = 3)
```


## AFDW-normalized metabolic rates
### GPP
```{r}
# linear model of log-transformed GPP (biomass-normalized) and log-transformed temperature

# linear model of GPP (AFDW-normalized)
log_GPP_AFDW_mod <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients*season, dat_pnld)
summary(log_GPP_AFDW_mod)

plot(log_GPP_AFDW_mod)
###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_GPP_AFDW_mod)["log(temp_C)"])


GPP_log_log <- dat_pnld %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = season)) +
  facet_wrap(~nutrient_trt) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

GPP_log_log


# investigating outlier datapoints using plotly to identify datapoints
GPP_plotly <- dat_pnld %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, color = season)) +
  facet_wrap(~nutrient_trt) +
  scale_y_log10() +
  geom_point(aes(group = chamber), shape = 19, size = 2) +
  geom_smooth(method = "lm") +
  #scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic()

p <- plotly::ggplotly(GPP_plotly)
p


# linear model of GPP (AFDW-normalized)
log_GPP_AFDW_mod <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients*season, dat_clean)
summary(log_GPP_AFDW_mod)

plot(log_GPP_AFDW_mod)
###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_GPP_AFDW_mod)["log(temp_C)"])


GPP_log_log <- dat_clean %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_gAFDW, group = interaction(season, nutrients))) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(aes(shape = season, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15, 21)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  #scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton GPP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

GPP_log_log

ggsave("plots/main_analysis/GPP_log_log_main.png", plot = GPP_log_log, width = 4, height = 2.5)
```


### ER
```{r}
# linear model of ER (AFDW-normalized)
log_ER_AFDW_mod <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients*season, dat_clean)
summary(log_ER_AFDW_mod)

plot(log_ER_AFDW_mod)

###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_ER_AFDW_mod)["log(temp_C)"])


ER_log_log <- dat_clean %>%
  ggplot(aes(x = temp_C, y = ER_mgO2_d_gAFDW, group = interaction(season, nutrients))) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point(aes(shape = season, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15, 21)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  #scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton ER vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))

ER_log_log

ggsave("plots/main_analysis/ER_log_log_main.png", plot = ER_log_log, width = 4, height = 2.5)
```

### NEP
```{r}
# linear model of NEP (AFDW-normalized)
log_NEP_AFDW_mod <- lm(log(NEP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients*season, dat_clean)
summary(log_NEP_AFDW_mod)

# plotting residuals by season
#boxplot(resid(log_NEP_AFDW_mod) ~ dat_clean)

plot(log_NEP_AFDW_mod)
###### Model interpretations ######

# interpreting effect size of temperature effect
exp(coef(log_NEP_AFDW_mod)["log(temp_C)"])


NEP_log_log <- dat_clean %>%
  ggplot(aes(x = temp_C, y = NEP_mgO2_d_gAFDW, group = interaction(season, nutrients))) +
  scale_y_log10() +
  scale_x_log10() +
 geom_point(aes(shape = season, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15, 21)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  #scale_color_discrete(name = "Nutrient Treatment")+
  theme_classic() +
  labs(title = "Periphyton NEP vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW^{-1})))

NEP_log_log

ggsave("plots/main_analysis/NEP_log_log_main.png", plot = NEP_log_log, width = 4, height = 2.5)
```

These linear models with no random effects may be overstating their confidence in the significance of the predictors since pooling the variance between seasons might be masking some of the season-level correlation (or covariance) between predictors like temperature.

# Faceting the above plots
```{r}
# melting dataframe to long format so that metabolic rate columns are combined into one column and values are in a separate column so all can be plotted at once
dat_long <- dat_clean %>%
  rename(GPP = GPP_mgO2_d_gAFDW, ER = ER_mgO2_d_gAFDW, NEP = NEP_mgO2_d_gAFDW) %>% # renaming columns so that facet labels are more tidy
  pivot_longer(cols = c(NEP, GPP, ER),
               names_to = "rate_name_AFDW",
               values_to = "rates_AFDW",
               values_drop_na = TRUE)
#View(dat_long)

# plotting same as above plots, but facetting by rate_name so that the temperature axis is shared between plots

rates_log_log <- dat_long %>%
  ggplot(aes(x = temp_C, y = rates_AFDW, group = interaction(season, nutrients))) +
  facet_grid(cols = vars(rate_name_AFDW)) +
  scale_y_log10() +
  scale_x_log10() +
 geom_point(aes(shape = season, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Season") +# renaming label of point shape legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "Periphyton Metabolic Rates vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~gAFDW^{-1})))

rates_log_log

ggsave("plots/main_analysis/rates_log_log_main.png", plot = rates_log_log, width = 7, height = 3)


# simplified plot for AGU 2021 poster
# making experiment ID column a factor instead of numeric since it is treated as discrete rather than continuous in the plot
dat_long$exp_ID <- as.factor(dat_long$exp_ID)

# now replotting the above plot with the title removed and exp_ID replacing season as a random effect in the model
rates_log_log_AGU <- dat_long %>%
  ggplot(aes(x = temp_C, y = rates_AFDW, group = interaction(exp_ID, nutrients))) +
  facet_grid(cols = vars(rate_name_AFDW)) +
  scale_y_log10() +
  scale_x_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Experiment ID") +# renaming label of point shape legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~gAFDW^{-1})))

rates_log_log_AGU

ggsave("plots/main_analysis/rates_log_log_AGU.png", plot = rates_log_log_AGU, width = 7, height = 3)

# high quality dpi plot
ggsave("plots/main_analysis/rates_log_log_AGU_hr.png", plot = rates_log_log_AGU, width = 7, height = 3, dpi = 300)
```

## Un-logged x-axis (temperature)
```{r}
rates_log_log_AGU <- dat_long %>%
  ggplot(aes(x = temp_C, y = rates_AFDW, group = interaction(exp_ID, nutrients))) +
  facet_grid(cols = vars(rate_name_AFDW)) +
  scale_y_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15, 21)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Experiment ID") +# renaming label of point shape legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~gAFDW^{-1})))

rates_log_log_AGU

# high quality dpi plot
ggsave("plots/main_analysis/rates_unlogged_AGU_hr.png", plot = rates_log_log_AGU, width = 7, height = 3, dpi = 300)

```


# New AFDW-adjusted rate plot
```{r}
# remelting dataframe to long format so that metabolic rate columns are combined into one column and values are in a separate column so all can be plotted at once
dat_long <- dat_clean %>%
  rename(GPP = GPP_mgO2_d_gAFDW_2, ER = ER_mgO2_d_gAFDW_2, NEP = NEP_mgO2_d_gAFDW_2) %>% # renaming columns so that facet labels are more tidy
  pivot_longer(cols = c(NEP, GPP, ER),
               names_to = "rate_name_AFDW",
               values_to = "rates_AFDW",
               values_drop_na = TRUE)
#View(dat_long)

# making experiment ID column a factor instead of numeric since it is treated as discrete rather than continuous in the plot
dat_long$exp_ID <- as.factor(dat_long$exp_ID)

AFDW_2_log_rates <- dat_long %>%
  ggplot(aes(x = temp_C, y = rates_AFDW, group = interaction(exp_ID, nutrients))) +
  facet_grid(cols = vars(rate_name_AFDW)) +
  scale_y_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Experiment ID") +# renaming label of point shape legend
  scale_shape_manual(values = c(0:6)) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~gAFDW^{-1})))

AFDW_2_log_rates

# high quality dpi plot
ggsave("plots/main_analysis/correct_AFDW_norm_rates.png", plot = AFDW_2_log_rates, width = 7, height = 3, dpi = 300)


# same plot, but ungrouping by experiment ID

metab_rates <- dat_long %>%
  ggplot(aes(x = temp_C, y = rates_AFDW, group = nutrients)) +
  facet_grid(cols = vars(rate_name_AFDW)) +
  scale_y_log10() +
 geom_point(aes(color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm", se = FALSE) +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~gAFDW^{-1})))

metab_rates

# high quality dpi plot
ggsave("plots/main_analysis/AFDW_model_output_plot.png", plot = metab_rates, width = 7, height = 3, dpi = 300)
```

# Plotting Areal Rates
```{r}
# melting dataframe to long format so that metabolic rate columns are combined into one column and values are in a separate column so all can be plotted at once
dat_long_m2 <- dat_clean %>%
  rename(GPP = GPP_mgO2_d_m2, ER = ER_mgO2_d_m2, NEP = NEP_mgO2_d_m2) %>% # renaming columns so that facet labels are more tidy
  pivot_longer(cols = c(NEP, GPP, ER),
               names_to = "rate_name_m2",
               values_to = "rates_m2",
               values_drop_na = TRUE)
#View(dat_long_m2)

# plotting same as above plots, but facetting by rate_name so that the temperature axis is shared between plots

areal_rates_log_log <- dat_long_m2 %>%
  ggplot(aes(x = temp_C, y = rates_m2, group = interaction(season, nutrients))) +
  facet_grid(cols = vars(rate_name_m2)) +
  scale_y_log10() +
  scale_x_log10() +
 geom_point(aes(shape = season, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15, 21)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Season") +# renaming label of point shape legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "Periphyton Areal Metabolic Rates vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~m^{-2})))

areal_rates_log_log

ggsave("plots/main_analysis/areal_rates_log_log_main.png", plot = areal_rates_log_log, width = 7, height = 3)


# simplified plot for AGU 2021 poster

# making experiment ID column a factor instead of numeric since it is treated as discrete rather than continuous in the plot
dat_long_m2$exp_ID <- as.factor(dat_long_m2$exp_ID)

# now replotting the above plot with the title removed and exp_ID replacing season as a random effect in the model
areal_rates_log_log_AGU <- dat_long_m2 %>%
  ggplot(aes(x = temp_C, y = rates_m2, group = interaction(exp_ID, nutrients))) +
  facet_grid(cols = vars(rate_name_m2)) +
  scale_y_log10() +
  scale_x_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15, 21)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Experiment ID") +# renaming label of point shape legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~m^{-2})))

areal_rates_log_log_AGU

ggsave("plots/main_analysis/areal_rates_log_log_AGU.png", plot = areal_rates_log_log_AGU, width = 7, height = 3, dpi = 300)
```

## Same plot with unlogged x-axis (temperature)
```{r}
areal_rates_log_log_AGU <- dat_long_m2 %>%
  ggplot(aes(x = temp_C, y = rates_m2, group = interaction(exp_ID, nutrients))) +
  facet_grid(cols = vars(rate_name_m2)) +
  scale_y_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Experiment ID") +# renaming label of point shape legend
  scale_shape_manual(values = c(0:6)) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~m^{-2})))

areal_rates_log_log_AGU

ggsave("plots/main_analysis/areal_rates_unlogged_AGU.png", plot = areal_rates_log_log_AGU, width = 7, height = 3, dpi = 300)


# same plot, but ungrouping by experiment ID

areal_rates <- dat_long_m2 %>%
  ggplot(aes(x = temp_C, y = rates_m2, group = nutrients)) +
  facet_grid(cols = vars(rate_name_m2)) +
  scale_y_log10() +
 geom_point(aes(color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm", se = FALSE) +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~m^{-2})))

areal_rates

ggsave("plots/main_analysis/SA_model_output.png", plot = areal_rates, width = 7, height = 3, dpi = 300)
```

# Chl-a-normalized GPP plot
```{r}
# normalizing GPP rates by mass of chla on rocks
chla_rates <- dat_clean %>%
  mutate(GPP_mgO2_d_ugChla = GPP_mg_d/(total_chla_ug/1e6)) #converting chla data from ug to g

# making experiment ID column a factor instead of numeric since it is treated as discrete rather than continuous in the plot
chla_rates$exp_ID <- as.factor(chla_rates$exp_ID)

# now plotting GPP rates vs. temp

GPP_chla <- chla_rates %>%
  ggplot(aes(x = temp_C, y = GPP_mgO2_d_ugChla, group = interaction(exp_ID, nutrients))) +
  scale_y_log10() +
 geom_point(aes(shape = exp_ID, color = nutrient_trt), size = 1.5) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Experiment ID") +# renaming label of point shape legend
  scale_shape_manual(values = c(0:6)) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gChla^{-1})))

GPP_chla

ggsave("plots/main_analysis/GPP_chla.png", plot = GPP_chla, height = 4, width = 6)

```


# Autotrophic Index
```{r}
# calculating the autotrophic index (ratio of AFDW:chl-a) (Biggs & Close 1989)

dat <- dat %>%
  mutate(AI = total_AFDW_g_2/(total_chla_ug/1e6)) # converting chla from ug to g so that the units cancel (g/g) in the AI ratio
         
dat %>%
  group_by(date) %>%
  summarise(AI_avg = mean(AI), AI_sd = sd(AI)) %>%
  arrange(AI_avg)

# plotting changes in AI over the year

AI_trends <- dat %>%
  filter(AI < 5000) %>% # filtering out 2 outliers from october and march
  ggplot(aes(x = date, y = AI, group = date)) +
  geom_boxplot() +
  labs(x = "Date", y = "Periphyton Autotrophic Index (gAFDW:gChla)") +
  theme_classic()

AI_trends

ggsave("plots/main_analysis/AI_timeseries.png", plot = AI_trends, height = 4, width = 8)
```


# Exploratory Data Analysis
```{r}
# histogram of temperature data to check normality
dat_pnld %>%
  ggplot(aes(temp_C)) +
  geom_histogram()
# looks approximately like a normal distribution

# now checking normality in AFDW-normalized metabolic rates
dat_pnld %>%
  ggplot(aes(GPP_mgO2_d_gAFDW)) +
  geom_histogram()

dat_pnld %>%
  ggplot(aes(NEP_mgO2_d_gAFDW)) +
  geom_histogram()

dat_pnld %>%
  ggplot(aes(ER_mgO2_d_gAFDW)) +
  geom_histogram()
# some definite outliers present, also some potential for a right-skewed distn in the ER and GPP rates.

# cleveland plot
# dotchart(dat_pnld$)

# selecting only the numeric columns for calculating a correlation matrix
dat_cor <- dat_pnld %>%
  dplyr::select(temp_C, GPP_mgO2_d_gAFDW, NEP_mgO2_d_gAFDW, ER_mgO2_d_gAFDW, GPP_mgO2_d_m2, NEP_mgO2_d_m2, ER_mgO2_d_m2, NO3_N_ug_L, NH4_N_ug_L, SRP_ug_L)

# standardizing data with mean = 0, sd = 1 for correlation matrix
dat_cor <- scale(dat_cor)
View(dat_cor)

# plotting correlation matrix of the data
cor(dat_cor, use = "complete.obs")

```



# Random Effects Modeling
## GPP
```{r}
# nested random effects model with temp and nutrients as predictors, and season as a random effect (with differing intercept and slope for each season)

# this basically is saying that the effect of increasing temperature varies among seasons, and that each season has different GPP values
m6 <- lmer(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1 + log(temp_C)|season), data = dat_clean)
display(m6)
summary (m6)


# random effects model with random intercept for season only

m5 <- lmer(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1|season), data = dat_clean)
display(m5)
summary(m5)

m7 <- lmer(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1 + nutrients|season), data = dat_clean)

m8 <- lmer(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1 + nutrients|season) + (1 + log(temp_C)|season), data = dat_clean)

m3 <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, data = dat_clean)

m1 <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), data = dat_clean)

m2 <- lm(log(GPP_mgO2_d_gAFDW) ~ nutrients, data = dat_clean)

m4 <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients*season, data = dat_clean)
summary(m8)

summary(m4)


AIC(m1, m2, m3, m4, m5, m6, m7, m8)
BIC(m1, m2, m3, m4, m5, m6, m7, m8)

```

Looks like model 8 has the lowest AIC and is the most parsimonious model with temperature, nutrients, season, and their interaction as predictors.

What does this model look like plotted out?

How to interpret significant 3-way interaction btwn temp, nuts, and season?

Does the interaction between temp and season even meaningful since we already know temp covaries with season?

## ER
```{r}
# nested random effects model with temp and nutrients as predictors, and season as a random effect (with differing intercept and slope for each season)

# this basically is saying that the effect of increasing temperature varies among seasons, and that each season has different ER values
m6 <- lmer(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1 + log(temp_C)|season), data = dat_clean)
display(m6)
summary (m6)


# random effects model with random intercept for season only

m5 <- lmer(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1|season), data = dat_clean)
display(m5)
summary(m5)

m7 <- lmer(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1 + nutrients|season), data = dat_clean)

m8 <- lmer(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1 + nutrients|season) + (1 + log(temp_C)|season), data = dat_clean)

m3 <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients, data = dat_clean)

m1 <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C), data = dat_clean)

m2 <- lm(log(ER_mgO2_d_gAFDW) ~ nutrients, data = dat_clean)

m4 <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients*season, data = dat_clean)
summary(m4)

# Best AIC-supported model
summary(m4)

AIC(m1, m2, m3, m4, m5, m6, m7, m8)
BIC(m1, m2, m3, m4, m5, m6, m7, m8)
```

# Random Effects W/o season
## GPP
```{r}
m1 <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), data = dat_clean)

m2 <- lm(log(GPP_mgO2_d_gAFDW) ~ nutrients, data = dat_clean)

m3 <- lmer(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) + (1|rock_ID), data = dat_clean)

m4 <- lmer(log(GPP_mgO2_d_gAFDW) ~ nutrients + (1|rock_ID), data = dat_clean)

m5 <- lmer(log(GPP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1|rock_ID), data = dat_clean)
display(m5)
summary (m5)

AIC(m1, m2, m3, m4, m5)
BIC(m1, m2, m3, m4, m5)
```

## ER
```{r}
m1 <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C), data = dat_clean)

m2 <- lm(log(ER_mgO2_d_gAFDW) ~ nutrients, data = dat_clean)

m3 <- lmer(log(ER_mgO2_d_gAFDW) ~ log(temp_C) + (1|rock_ID), data = dat_clean)

m4 <- lmer(log(ER_mgO2_d_gAFDW) ~ nutrients + (1|rock_ID), data = dat_clean)

m5 <- lmer(log(ER_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1|rock_ID), data = dat_clean)
display(m5)
summary (m5)

AIC(m1, m2, m3, m4, m5)
BIC(m1, m2, m3, m4, m5)
```

## NEP
```{r}
m1 <- lm(log(NEP_mgO2_d_gAFDW) ~ log(temp_C), data = dat_clean)

m2 <- lm(log(NEP_mgO2_d_gAFDW) ~ nutrients, data = dat_clean)

m3 <- lmer(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) + (1|rock_ID), data = dat_clean)

m4 <- lmer(log(NEP_mgO2_d_gAFDW) ~ nutrients + (1|rock_ID), data = dat_clean)

m5 <- lmer(log(NEP_mgO2_d_gAFDW) ~ log(temp_C)*nutrients + (1|rock_ID), data = dat_clean)
display(m5)
summary (m5)

AIC(m1, m2, m3, m4, m5)
BIC(m1, m2, m3, m4, m5)
```

# nlme models
## Initial method
### GPP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(GPP_mgO2_d_gAFDW))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

m1 <- lm(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), data = dat_lme)

m2 <- lm(log(GPP_mgO2_d_gAFDW) ~ nutrients, data = dat_lme)

m3 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, data = dat_lme)

m4 <- lme(log(GPP_mgO2_d_gAFDW) ~ nutrients, random = ~ 1 | rock_ID, data = dat_lme)

m5 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, data = dat_lme)

#m6 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C) | rock_ID, data = dat_lme)
# error message for this model saying it never converged

m7 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + nutrients | rock_ID, data = dat_lme)

#m8 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = list(~ 1 + nutrients | rock_ID, ~ 1 + log(temp_C) | rock_ID), data = dat_lme)
# error message for this model saying it did not converge

# incorporating weights to account for heteroscedasticity within and btwn different experiments
m9 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

# random intercept model for rock effect
m10 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

#m11 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + nutrients | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# would not converge

#m12 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = list(~ 1 | rock_ID, ~ 1 + temp_C | exp_ID), weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# does not converge

#m13 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = list(~ 1 | rock_ID, ~ 1 + nutrients | exp_ID), weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# does not converge

#m14 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = list(~ 1 | rock_ID, ~ 1 + temp_C | exp_ID, ~ 1 + nutrients | exp_ID), weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# does not converge

m15 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

# letting the model variance parameter vary by nutrient treatment and letting this vary between experiment ID. Needed to use VarComb so that I could use the VarExp option since the nutrient treatment contains values of "0" and I needed to combine this with letting sigma_sq vary across different experiments.
m16 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, weights = varComb(varIdent(form = ~1|exp_ID), varExp(form = ~ nutrients)), data = dat_lme)

m17 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, weights = varIdent(form = ~log(temp_C)|exp_ID), data = dat_lme)
# iteration limit reached without convergence

m18 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~log(temp_C)|exp_ID), data = dat_lme)
# false convergence

AIC(m1, m2, m3, m4, m5, m7, m9, m10, m15, m16, m17, m18)
BIC(m1, m2, m3, m4, m5, m7, m9, m10, m15, m16, m17, m18)

# looks like m10, the random intercept model for the rock effect, is the best model. Now lets look at which fixed effects are signficant
summary(m10)
# looks like only log(temp_C) is the only signifcant fixed effect, thus we can drop nutrients and their interaction.

mfinal <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

summary(mfinal)


# graphical model validation using standardized residuals vs. fitten values to look for heteroscedasticity
plot(m10)

```


## ER
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(ER_mgO2_d_gAFDW))
glimpse(dat_lme)

m1 <- lm(log(ER_mgO2_d_gAFDW) ~ log(temp_C), data = dat_lme)

m2 <- lm(log(ER_mgO2_d_gAFDW) ~ nutrients, data = dat_lme)

m3 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, data = dat_lme)

m4 <- lme(log(ER_mgO2_d_gAFDW) ~ nutrients, random = ~ 1 | rock_ID, data = dat_lme)

m5 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, data = dat_lme)

#m6 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C) | rock_ID, data = dat_lme)
# error message for this model saying it never converged

m7 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + nutrients | rock_ID, data = dat_lme)

#m8 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = list(~ 1 +  nutrients | rock_ID, ~ 1 + log(temp_C) | rock_ID), data = dat_lme)
# error message for this model saying it did not converge

# incorporating weights to account for heteroscedasticity within and btwn different experiments
m9 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

#m10 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + nutrients | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# would not converge


AIC(m1, m2, m3, m4, m5, m7, m9)
BIC(m1, m2, m3, m4, m5, m7, m9)

summary(m5)
```



## NEP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(NEP_mgO2_d_gAFDW))
glimpse(dat_lme)

m1 <- lm(log(NEP_mgO2_d_gAFDW) ~ log(temp_C), data = dat_lme)

m2 <- lm(log(NEP_mgO2_d_gAFDW) ~ nutrients, data = dat_lme)

m3 <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, data = dat_lme)

m4 <- lme(log(NEP_mgO2_d_gAFDW) ~ nutrients, random = ~ 1 | rock_ID, data = dat_lme)

m5 <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, data = dat_lme)

#m6 <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C) | rock_ID, data = dat_lme)
# error message for this model saying it never converged

m7 <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + nutrients | rock_ID, data = dat_lme)

m8 <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = list(~1 + nutrients | rock_ID, ~1 + log(temp_C) | rock_ID), data = dat_lme)

# incorporating weights to account for heteroscedasticity within and btwn experiments
#m9 <- 


AIC(m1, m2, m3, m4, m5, m7, m8)
BIC(m1, m2, m3, m4, m5, m7, m8)

m3
```

# Zuur method
### GPP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(GPP_mgO2_d_gAFDW))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, method = "REML", data = dat_lme)

m2 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding in random intercept for each rock sample

m3 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme) # incorporating weights to account for heteroscedasticity within and btwn different experiments

#m4 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C) | rock_ID, data = dat_lme, control=(msMaxIter=500)) # adding in random intercept and slope for each rock sample, also increasing the number of iterations to acheive model convergence
# would not converge

 #m5 <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C) | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500)) # incorporating weights to account for heteroscedasticity within and btwn different experiments, also increasing the number of iterations to acheive model convergence
# would not converge

AIC(m1, m2, m3)
# model 3 has the lowest AIC

summary(m3)
# the model summary indicates that the nutrients and temp:nutrients interaction are not significant, thus I will drop them from the model

GPP_final <- lme(log(GPP_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

summary(GPP_final)
```


#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(GPP_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(GPP_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```

### ER
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(ER_mgO2_d_gAFDW))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, method = "REML", data = dat_lme)

m2 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding in random intercept for each rock sample

m3 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

m4 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C) | rock_ID, data = dat_lme) # adding random intercept for rock and random slope for temperature

m5 <- lme(log(ER_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C) | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

AIC(m1, m2, m3, m4, m5)
# model 2 has the strongest support according to AIC

summary(m2)
# looks like all the fixed effects are statistically significant and will be kept in the model

ER_final <- m2

```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(ER_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(ER_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```


### NEP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(NEP_mgO2_d_gAFDW))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, method = "REML", data = dat_lme)

m2 <-  lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding random intercept for each rock

m3 <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme) # accounting for heteroscedasticity within and among different experiments

#m4 <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C)| rock_ID, data = dat_lme, control=(msMaxIter=500)) # adding random slopes for each rock (i.e. the relationship between temperature and NEP varies for each individual rock)
# model does not converge

#m5 <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C) * nutrients, random = ~ 1 + log(temp_C)| rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500)) # now adding in variance structure back in to account for heterscedasticity within and between experiments
# model does not converge

AIC(m1, m2, m3)
# model 3 has the most support according to AIC

summary(m3)
# the nutrient and temp:nutrient interaction fixed effects are not statistically significant so they will be dropped from the final model

NEP_final <- lme(log(NEP_mgO2_d_gAFDW) ~ log(temp_C), random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

summary(NEP_final)

```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(NEP_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(NEP_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```


# Zuur method with un-transformed temp predictor
## GPP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(GPP_mgO2_d_gAFDW))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(GPP_mgO2_d_gAFDW) ~ temp_C * nutrients, method = "REML", data = dat_lme)

m2 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding in random intercept for each rock sample

m3 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme) # incorporating weights to account for heteroscedasticity within and btwn different experiments

#m4 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 + temp_C | rock_ID, data = dat_lme, control=(msMaxIter=500)) # adding in random intercept and slope for each rock sample, also increasing the number of iterations to acheive model convergence
# would not converge

m5 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 + temp_C | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500)) # incorporating weights to account for heteroscedasticity within and btwn different experiments, also increasing the number of iterations to acheive model convergence

AIC(m1, m2, m3, m5)
# model 5 has the lowest AIC

summary(m5)
# the model summary indicates that the nutrients and temp:nutrients interaction are not significant, thus I will drop them from the model

GPP_final <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_C, random = ~ 1 + temp_C | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500))

summary(GPP_final)

# interpreting effect sizes of temp
exp(0.0383)
# for every 1C increase in temperature, GPP increases by 4% (a factor of 1.039 mgO2 per day per g AFDW).
```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(GPP_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(GPP_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```

## ER
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(ER_mgO2_d_gAFDW))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(ER_mgO2_d_gAFDW) ~ temp_C * nutrients, method = "REML", data = dat_lme)

m2 <- lme(log(ER_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding in random intercept for each rock sample

m3 <- lme(log(ER_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

#m4 <- lme(log(ER_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 + temp_C | rock_ID, data = dat_lme) # adding random intercept for rock and random slope for temperature
# does not converge

#m5 <- lme(log(ER_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 + temp_C | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# does not converge

AIC(m1, m2, m3)
# model 3 has the strongest support according to AIC

summary(m3)
# looks like all the fixed effects are statistically significant and will be kept in the model

ER_final <- m3

# interpreting effect size of temp
exp(0.0658)

# for every 1C increase in temperature, ER increases by 7% (a factor of 1.068 mg O2 per day per g AFDW)

# interpreting effect size of nutrients
exp(0.52)
```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(ER_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(ER_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```

## NEP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(NEP_mgO2_d_gAFDW))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(NEP_mgO2_d_gAFDW) ~ temp_C * nutrients, method = "REML", data = dat_lme)

m2 <-  lme(log(NEP_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding random intercept for each rock

m3 <- lme(log(NEP_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme) # accounting for heteroscedasticity within and among different experiments

#m4 <- lme(log(NEP_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 + temp_C| rock_ID, data = dat_lme, control=(msMaxIter=500)) # adding random slopes for each rock (i.e. the relationship between temperature and NEP varies for each individual rock)
# model does not converge

m5 <- lme(log(NEP_mgO2_d_gAFDW) ~ temp_C * nutrients, random = ~ 1 + temp_C| rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500)) # now adding in variance structure back in to account for heterscedasticity within and between experiments
# model does not converge

AIC(m1, m2, m3, m5)
# model 5 has the most support according to AIC

summary(m5)
# the nutrient and temp:nutrient interaction fixed effects are not statistically significant so they will be dropped from the final model

NEP_final <- lme(log(NEP_mgO2_d_gAFDW) ~ temp_C, random = ~ 1 + temp_C| rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500))

summary(NEP_final)

# interpreting effect size of temp
exp(0.0376)

# for every 1C increase in temperature, NEP increases by 3.8% (a factor of 1.038 mg O2 per day per g AFDW)
```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(NEP_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(NEP_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```


# New AFDW + Zuur method with un-transformed temp predictor
### GPP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(GPP_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(GPP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, method = "REML", data = dat_lme)

m2 <- lme(log(GPP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding in random intercept for each rock sample

m3 <- lme(log(GPP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme) # incorporating weights to account for heteroscedasticity within and btwn different experiments

m4 <- lme(log(GPP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 + temp_C | rock_ID, data = dat_lme, control=(msMaxIter=500)) # adding in random intercept and slope for each rock sample, also increasing the number of iterations to acheive model convergence


#m5 <- lme(log(GPP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 + temp_C | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=1000)) # incorporating weights to account for heteroscedasticity within and btwn different experiments, also increasing the number of iterations to acheive model convergence
# would not converge

AIC(m1, m2, m3, m4)
# model 3 has the lowest AIC

summary(m3)
# the model summary indicates that the nutrients and temp:nutrients interaction are not significant, thus I will drop them from the model

GPP_final <- lme(log(GPP_mgO2_d_gAFDW_2) ~ temp_C, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

summary(GPP_final)

# interpreting effect sizes of temp
exp(0.0422)
# for every 1C increase in temperature, GPP increases by 4.3% (a factor of 1.043 mgO2 per day per g AFDW).
```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(GPP_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(GPP_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```

### ER
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(ER_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(ER_mgO2_d_gAFDW_2) ~ temp_C * nutrients, method = "REML", data = dat_lme)

m2 <- lme(log(ER_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding in random intercept for each rock sample

m3 <- lme(log(ER_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

m4 <- lme(log(ER_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 + temp_C | rock_ID, data = dat_lme) # adding random intercept for rock and random slope for temperature


#m5 <- lme(log(ER_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 + temp_C | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# does not converge

AIC(m1, m2, m3, m4)
# model 3 has the strongest support according to AIC

summary(m3)
# looks like all the fixed effects are statistically significant and will be kept in the model

ER_final <- m3

# interpreting effect size of temp
exp(0.0660)

# for every 1C increase in temperature, ER increases by 7% (a factor of 1.068 mg O2 per day per g AFDW)

# interpreting effect size of nutrients
exp(0.496)

# enriched ER rates were increased by 64% on average relative to ambient rates

# interpreting temp:nutrient interaction effect
exp(-.0169)
```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(ER_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(ER_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```

### NEP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(NEP_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

glimpse(dat_lme)
#View(dat_lme)

# starting with as many explanatory variables as possible in the fixed component of the models (temp, nutrients, and their interaction)
m1 <- gls(log(NEP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, method = "REML", data = dat_lme)

m2 <-  lme(log(NEP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 | rock_ID, data = dat_lme) # adding random intercept for each rock

m3 <- lme(log(NEP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme) # accounting for heteroscedasticity within and among different experiments

m4 <- lme(log(NEP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 + temp_C| rock_ID, data = dat_lme, control=(msMaxIter=500)) # adding random slopes for each rock (i.e. the relationship between temperature and NEP varies for each individual rock)
# model does not converge

m5 <- lme(log(NEP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 + temp_C| rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500)) # now adding in variance structure back in to account for heterscedasticity within and between experiments
# model does not converge

AIC(m1, m2, m3, m4, m5)
# model 5 has the most support according to AIC

summary(m5)
# the nutrient and temp:nutrient interaction fixed effects are not statistically significant so they will be dropped from the final model

NEP_final <- lme(log(NEP_mgO2_d_gAFDW_2) ~ temp_C, random = ~ 1 + temp_C| rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500))

summary(NEP_final)

# interpreting effect size of temp
exp(0.0351)

# for every 1C increase in temperature, NEP increases by 3.6% (a factor of 1.036 mg O2 per day per g AFDW)
```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(NEP_final)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(NEP_final, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```

Modeling based on corrected AFDW values did not affect the results or interpretation of the models drastically, and some effect sizes for the fixed effects changed minimally.

### Plotting Model Results
#### GPP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(GPP_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

# saving effect size estimates into a dataframe
library(effects)
effects_temp <- effects::effect(term = "temp_C", mod = GPP_final)
summary(effects_temp)

# saving effects in a dataframe
temp_df <- as.data.frame(effects_temp)
head(temp_df)

GPP_mod_plot <- ggplot() +
  scale_y_log10() +
  geom_point(data = dat_lme, aes(x = temp_C, y = GPP_mgO2_d_gAFDW_2),size = 1.5) +
  geom_line(data = temp_df, aes(x = temp_C, y = exp(fit))) + # un-logging estimates so that they are plotted with the un-logged data on a log10 axis
  geom_ribbon(data = temp_df, aes(x = temp_C, ymin = exp(lower), ymax = exp(upper)), alpha = 0.3) +
  theme_classic() +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(GPP~(mgO[2]~day^{-1}~gAFDW^{-1})))

GPP_mod_plot
```

#### ER
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(ER_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)


# saving effect size estimates into a dataframe
library(effects)
effects_temp <- effects::effect(term = "temp_C", mod = ER_final)
summary(effects_temp)

# saving effects in a dataframe
temp_df <- as.data.frame(effects_temp)
head(temp_df)

ER_mod_plot <- ggplot() +
  scale_y_log10() +
  geom_point(data = dat_lme, aes(x = temp_C, y = ER_mgO2_d_gAFDW_2),size = 1.5) +
  geom_line(data = temp_df, aes(x = temp_C, y = exp(fit))) + # un-logging estimates so that they are plotted with the un-logged data on a log10 axis
  geom_ribbon(data = temp_df, aes(x = temp_C, ymin = exp(lower), ymax = exp(upper)), alpha = 0.3) +
  theme_classic() +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(ER~(mgO[2]~day^{-1}~gAFDW^{-1})))

ER_mod_plot
```


#### NEP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(NEP_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)


# saving effect size estimates into a dataframe
library(effects)
effects_temp <- effects::effect(term = "temp_C", mod = NEP_final)
summary(effects_temp)

# saving effects in a dataframe
temp_df <- as.data.frame(effects_temp)
head(temp_df)

NEP_mod_plot <- ggplot() +
  scale_y_log10() +
  geom_point(data = dat_lme, aes(x = temp_C, y = NEP_mgO2_d_gAFDW_2),size = 1.5) +
  geom_line(data = temp_df, aes(x = temp_C, y = exp(fit))) + # un-logging estimates so that they are plotted with the un-logged data on a log10 axis
  geom_ribbon(data = temp_df, aes(x = temp_C, ymin = exp(lower), ymax = exp(upper)), alpha = 0.3) +
  theme_classic() +
  labs(x = expression('Temperature (\u00B0C)'),
      y = expression(NEP~(mgO[2]~day^{-1}~gAFDW^{-1})))

NEP_mod_plot
```


# Final models

Using advice from Andrew Blandino of UCD stats consulting, the following models will keep all fixed effects as these are inherent in the experimental design and thus should not be left out of the model despite statistical insignificance.

## GPP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(GPP_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

# model with lowest AIC from GPP analyis above
m3 <- lme(log(GPP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 | rock_ID / exp_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme) # incorporating weights to account for heteroscedasticity within and btwn different experiments, also allowing random intercepts for each rock and exp_ID combo

summary(m3)
```
#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(m3)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(m3, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```


## ER
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(ER_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

# model 3 has lowest AIC from ER analysis above
m3 <- lme(log(ER_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 | rock_ID/exp_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme) # allowing random slope term to vary by rock IDs and experiment ID (april 2021 experiment did not have repeated measures while subsequent ones did)

summary(m3)
```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(m3)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(m3, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```

## NEP
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_clean %>%
  filter(!is.na(NEP_mgO2_d_gAFDW_2))

# converting experiment ID from numberic to a factor
dat_lme$exp_ID <- as.factor(dat_lme$exp_ID)

# m5 has the lowest AIC from the NEP analysis above
m5 <- lme(log(NEP_mgO2_d_gAFDW_2) ~ temp_C * nutrients, random = ~ 1 + temp_C| rock_ID/exp_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme, control=(msMaxIter=500)) # allowing intercept to vary by rock ID and experiment ID

summary(m5)
```

#### Graphical model validation
```{r}
# checking ordinary residuals of the final model (these are allowed to show heteroscedasticity). Using code from Zuur et al.
E1 <- resid(m5)
coplot(E1 ∼ rock_ID|exp_ID,
ylab = "Ordinary residuals", data = dat_lme)

# now checking standardized residuals of each of the experiments (random effect in the model). These should not show any patterns and should exhibit homogeneous variance.
E2 <- resid(m5, type = "normalized")
coplot(E2 ∼ rock_ID|exp_ID,
ylab = "Normalized residuals", data = dat_lme)

#coplot(resid(m10) ~ fitted(m10) | dat_lme$exp_ID)
```


# Centering Temp and Nutrient Data by experiment
```{r}
# centering temperature variable to "de-seasonalize" it since baseline temperatures change between seasons. This way we get can model the "treatment condition itself" rather than season. - stats consulting dept in UC Davis

# creating a column of average temperatures for each experiment
dat_centered <- dat_clean %>%
  group_by(month) %>%
  summarise(avg_T = mean(temp_C))

# adding avg T column to main dataframe so that we can mutate a new column of centered temps with it
dat_centered <- left_join(dat_clean, dat_centered, by = "month")

head(dat_centered)

# now subtracting average temp for each incubation from the actual temp to center this variable and remove the "seasonal effect" from temp data
dat_centered <- dat_centered %>%
  mutate(temp_centered = temp_C - avg_T)

head(dat_centered)
#View(dat_centered)
```

##### Plotting centered temp data
```{r}
# first plotting up data to see what it looks like

# melting dataframe to long format so that metabolic rate columns are combined into one column and values are in a separate column so all can be plotted at once
dat_long <- dat_centered %>%
  rename(GPP = GPP_mgO2_d_gAFDW, ER = ER_mgO2_d_gAFDW, NEP = NEP_mgO2_d_gAFDW) %>% # renaming columns so that facet labels are more tidy
  pivot_longer(cols = c(NEP, GPP, ER),
               names_to = "rate_name_AFDW",
               values_to = "rates_AFDW",
               values_drop_na = TRUE)
#View(dat_long)
rates_log_log <- dat_long %>%
  ggplot(aes(x = temp_centered, y = rates_AFDW, group = interaction(season, nutrients))) +
  facet_grid(cols = vars(rate_name_AFDW)) +
  scale_y_log10() +
 geom_point(aes(shape = season, color = nutrient_trt), size = 1.5) +
  scale_shape_manual(values = c(17, 19, 15)) +
  geom_smooth(aes(color = nutrient_trt), method = "lm") +
  scale_color_viridis_d("Nutrient Treatment") + # renaming label of color legend
  scale_shape_discrete(name = "Season") +# renaming label of point shape legend
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "Periphyton Metabolic Rates vs. Temperature",
      x = expression('Temperature (\u00B0C)'),
      y = expression(Rates~(mgO[2]~day^{-1}~gAFDW^{-1})))

rates_log_log

```

##### Modeling rates with centered temp data
```{r}
# making new dataset for lme models to remove NAs
dat_lme <- dat_centered %>%
  filter(!is.na(GPP_mgO2_d_gAFDW))
glimpse(dat_lme)

m1 <- lm(log(GPP_mgO2_d_gAFDW) ~ temp_centered, data = dat_lme)

m2 <- lm(log(GPP_mgO2_d_gAFDW) ~ nutrients, data = dat_lme)

m3 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered, random = ~ 1 | rock_ID, data = dat_lme)

m4 <- lme(log(GPP_mgO2_d_gAFDW) ~ nutrients, random = ~ 1 | rock_ID, data = dat_lme)

m5 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = ~ 1 | rock_ID, data = dat_lme)

#m6 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = ~ 1 + temp_centered | rock_ID, data = dat_lme)
# error message for this model saying it never converged

m7 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = ~ 1 + nutrients | rock_ID, data = dat_lme)

#m8 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = list(~ 1 + nutrients | rock_ID, ~ 1 + temp_centered | rock_ID), data = dat_lme)
# error message for this model saying it did not converge

# incorporating weights to account for heteroscedasticity within and btwn different experiments
m9 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

m10 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

#m11 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = ~ 1 + nutrients | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# would not converge

#m12 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = list(~ 1 | rock_ID, ~ 1 + temp_centered | exp_ID), weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# does not converge

#m13 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = list(~ 1 | rock_ID, ~ 1 + nutrients | exp_ID), weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# does not converge

#m14 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = list(~ 1 | rock_ID, ~ 1 + temp_centered | exp_ID, ~ 1 + nutrients | exp_ID), weights = varIdent(form = ~1|exp_ID), data = dat_lme)
# does not converge

m15 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered, random = ~ 1 | rock_ID, weights = varIdent(form = ~1|exp_ID), data = dat_lme)

# letting the model variance parameter vary by nutrient treatment and letting this vary between experiment ID. Needed to use VarComb so that I could use the VarExp option since the nutrient treatment contains values of "0" and I needed to combine this with letting sigma_sq vary across different experiments.
m16 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = ~ 1 | rock_ID, weights = varComb(varIdent(form = ~1|exp_ID), varExp(form = ~ nutrients)), data = dat_lme)


# letting the model variance parameter vary by temperature treatment and letting this vary between experiment ID
m17 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered, random = ~ 1 | rock_ID, weights = varIdent(form = ~temp_centered|exp_ID), data = dat_lme)


m18 <- lme(log(GPP_mgO2_d_gAFDW) ~ temp_centered * nutrients, random = ~ 1 | rock_ID, weights = varIdent(form = ~temp_centered|exp_ID), data = dat_lme)


AIC(m1, m2, m3, m4, m5, m7, m9, m10, m15, m16, m17, m18)
BIC(m1, m2, m3, m4, m5, m7, m9, m10, m15, m16, m17, m18)

m3
```



