---
title: "Waterbath Temp Tests"
author: "Nick Framsted"
date: "12/8/2020"
output: pdf_document
---
# Summary: 
This script contains analysis of temperature tests of newly fabricated temperature-controlled waterbaths from Oceans Design to be used for incubations of periphyton for the Lahontan Regional Water Quality Board. Below are graphs of tests ran on 3 separate dates testing the effects of overhead lighting on heating/cooling capacity and the speed at which the waterbaths (and the water-filled incubation chambers they contain) heat and cool under different conditions. 

#Quick conclusions:
Waterbaths cool ~2.7-3.3C/hr, regardless of lights on or off (nov 19 & 23 exps.). Chambers tend to heat/cool at nearly half the rate of the water baths ~1.35C/hr (nov 23 exp.), which can mean for longer equilibration times but also means greater temperature stability at a given temperature set point. Waterbaths heat ~0.5C/hr just with lights on, thus overhead lightling did not have a large effect on temperature stability (nov 19 & 23 exps.). However tank heating/chilling efficiency was very low when operating at low temps (Nov 24th exp.).

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(ggpubr)
# changing working directory from /scripts to /tahoe_lab_inc_analysis
knitr::opts_knit$set(root.dir = '..')
```

# reading in data
```{r, include=FALSE}
getwd()
dat <- read_csv("data/waterbath_temptests_main.csv", skip = 1)
```


# converting date-times in data
```{r}
str(dat)
head(dat)
dat$date_time_PST <- mdy_hms(dat$date_time_GMT_0800, tz = "America/Los_Angeles")
head(dat)

# and adding in day column
dat$day <- day(dat$date_time_PST)

head(dat)

#making tank a factor instead of numeric (it's really a categorical variable)
dat$tank <- as.factor(dat$tank)
```

# Nov 19 experiment
```{r}
# filtering by date time to only include measurements taken during nov 19th and trimming off data taken before start time and after end time of the experiment 
dat_19 <- dat %>%
  filter(date_time_PST > "2020-11-19 14:45:00" & date_time_PST < "2020-11-19 16:30:00")
dat_19 %>%
  ggplot(aes(x = date_time_PST, y = temp_C, color = tank)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_bw()

# cooling period only
dat_19 <- dat %>%
  filter(date_time_PST > "2020-11-19 14:45:00" & date_time_PST < "2020-11-19 15:30:00")
dat_19 %>%
  ggplot(aes(x = date_time_PST, y = temp_C, color = tank)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_bw()

# warming period
  dat_19 <- dat %>%
  filter(date_time_PST > "2020-11-19 15:45:00" & date_time_PST < "2020-11-19 16:25:00")
dat_19 %>%
  ggplot(aes(x = date_time_PST, y = temp_C, color = tank)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_bw()
  
  
```
# conclusions
There is no real difference in cooling when you have lights on or when they are off. Lights on will warm tanks about 1.7 times faster when the heat pump is disengaged. 

# Nov 23 experiment
```{r}
# filtering by date time to only include measurements taken during nov 23th and trimming data, only doing tanks 1 & 3 to analyze light effects on heating.
# tank 1 = lights on, tank 3 = lights off
dat_23 <- dat %>%
  filter(date_time_PST > "2020-11-23 14:00:00" & date_time_PST < "2020-11-23 16:30:00")
dat_23 %>%
  ggplot(aes(x = date_time_PST, y = temp_C, color = container)) +
  facet_wrap(~tank) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_bw()


```

# tanks 1 & 3 only
```{r}
# only doing tanks 1 & 3 to analyze light effects on heating
dat_23 <- dat %>%
  filter(date_time_PST > "2020-11-23 14:00:00" & date_time_PST < "2020-11-23 16:30:00") %>%
  filter(tank == 1 | tank == 3)
dat_23 %>%
  ggplot(aes(x = date_time_PST, y = temp_C, color = container)) +
  facet_wrap(~tank) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_bw()
```


Looks like the tanks heated rapidly (~2C in 30mins) and then the heater/chiller shut off. The tank was then heated slowly by ambient temps

# tanks 2 & 4 only
```{r}
# only doing tanks 2 & 4 to analyze chamber effects on cooling
dat_23 <- dat %>%
  filter(date_time_PST > "2020-11-23 15:05:00" & date_time_PST < "2020-11-23 16:30:00") %>%
  filter(tank == 2 | tank == 4)
dat_23 %>%
  ggplot(aes(x = date_time_PST, y = temp_C, color = container)) +
  facet_wrap(~tank) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_bw()
```

Water in chambers cools at almost half the rate of the water bath, this could matter when making drastic changes to temperature.

# 24 Nov experiment
```{r}
# filtering by date time to only include measurements taken during nov 24th and trimming data
dat_24 <- dat %>%
  filter(date_time_PST > "2020-11-24 12:50:00")
dat_24 %>%
  ggplot(aes(x = date_time_PST, y = temp_C, color = container)) +
  facet_wrap(~tank) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation() +
  theme_bw()
```

This tank cooled much slower than expected, it was operating near it's minimum operating temperature and may have been less efficient, but the water bath never reached the 8.5C setpoint.